Dominance Regions for Rank Two Cluster Algebras

### 2022-04-08 dominance_regions.tex
Inequalities for negative mutations

  \begin{lemma}
    For $k\ge1$ and $\lambda\in\cI$, the region $\phi_k^{-1}\cC_k(\phi_k\lambda)\subset\RR^2$ is the polytope determined by the following inequalities:
    \begin{align}
      \label{ineq:1} u_k^-\mu_0+u_{k-1}^+\mu_1 &\le u_k^-\lambda_0+u_{k-1}^+\lambda_1;\\
      \label{ineq:2} u_{k+1}^-\mu_0+u_k^+\mu_1 &\le u_{k+1}^-\lambda_0+u_k^+\lambda_1;\\
      \label{ineq:3} -u_{k-1}^-\mu_0+u_k^+\mu_1 &\le u_{k+1}^-\lambda_0+u_k^+\lambda_1;\\
      \label{ineq:4} -u_{k-1}^-\mu_0-u_{k-2}^+\mu_1 &\le u_{k+1}^-\lambda_0+u_k^+\lambda_1.
    \end{align}

@@

    For $k=2j$, $j>0$, and $\lambda\in\cI$, we observe that $\cC_k(\phi_k\lambda)\subset\RR^2$ is given by the inequalities 
    \[ (\dagger)\ x\le u_{2j+1}^+\lambda_0+u_{2j}^+\lambda_1 \qquad\text{and}\qquad (\ddagger)\ -y\le u_{2j}^-\lambda_0+u_{2j-1}^-\lambda_1. \]
    We inductively compute the region $\phi_{2j}^{-1}\cC_{2j}(\phi_{2j}\lambda)$, using Lemma~\ref{le:transformed inequalities} and the equality $\phi_{2j}^{-1}=\phi_2^{-j}$.

@@

    For $k=2j+1$, $j>0$, and $\lambda\in\cI$, we get $\cC_{2j+1}(\phi_{2j+1}\lambda)\subset\RR^2$ is given by the inequalities 
    \[ (\dagger')\ -x\le u_{2j+1}^+\lambda_0+u_{2j}^+\lambda_1 \qquad\text{and}\qquad (\ddagger')\ y\le u_{2j+2}^-\lambda_0+u_{2j+1}^-\lambda_1.\]
    Using that $\phi_{2j+1}^{-1}=\phi_2^{-j}\phi_1^{-1}$, we compute the image inductively as above.
    From \eqref{eq:forward mutation 1} and Lemma~\ref{le:imaginary stability}, we see that the boundary ray for $\cC_{2j+1}(\phi_{2j+1}\lambda)$ corresponding to ($\dagger'$) lies entirely in the region in which $\phi_1^{-1}$ acts according to the matrix $\left[ \begin{array}{cc} -1 & 0\\ c & 1 \end{array}\right]$ of determinant $-1$.
    Thus applying Lemma~\ref{le:transformed inequalities}, the inequality ($\dagger'$) is transformed by $\phi_1^{-1}$ into the inequality $x\le u_{2j+1}^+\lambda_0+u_{2j}^+\lambda_1$.
    The boundary ray corresponding to ($\ddagger'$) intersects both domains of linearity for $\phi_1^{-1}$ and thus produces the inequalities
    \[ cx+y\le u_{2j+2}^+\lambda_0+u_{2j+1}^+\lambda_1 \qquad\text{and}\qquad y\le u_{2j+2}^+\lambda_0+u_{2j+1}^+\lambda_1.\]

@@

>>>>

  \begin{lemma}
    For $k\in\ZZ$, $k\ne0$, and $\lambda\in\cI$, the region $\phi_k^{-1}\cC_k(\phi_k\lambda)\subset\RR^2$ is the polytope determined by the following inequalities:
    \begin{align}
      \label{ineq:1} u_k^-\mu_0+u_{k-1}^+\mu_1 &\le u_k^-\lambda_0+u_{k-1}^+\lambda_1;\\
      \label{ineq:2} u_{k+1}^-\mu_0+u_k^+\mu_1 &\le u_{k+1}^-\lambda_0+u_k^+\lambda_1;\\
      \label{ineq:3} -u_{k-1}^-\mu_0+u_k^+\mu_1 &\le u_{k+1}^-\lambda_0+u_k^+\lambda_1;\\
      \label{ineq:4} -u_{k-1}^-\mu_0-u_{k-2}^+\mu_1 &\le u_{k+1}^-\lambda_0+u_k^+\lambda_1.
    \end{align}
    when $k\ge1$ and by the following inequalities:
    \begin{align*}
      u_{k+1}^-\mu_0+u_k^+\mu_1 &\le u_{k+1}^-\lambda_0+u_k^+\lambda_1;\\
      u_k^-\mu_0+u_{k-1}^+\mu_1 &\le u_k^-\lambda_0+u_{k-1}^+\lambda_1;\\
      u_k^-\mu_0-u_{k+1}^+\mu_1 &\le u_k^-\lambda_0+u_{k-1}^+\lambda_1;\\
      -u_{k+2}^-\mu_0-u_{k+1}^+\mu_1 &\le u_k^-\lambda_0+u_{k-1}^+\lambda_1.
    \end{align*}
    when $k\le -1$.

@@

    For $k=2j$, $j>0$, and $\lambda\in\cI$, we observe that $\cC_k(\phi_k\lambda)\subset\RR^2$ is given by the inequalities 
    \[ (\dagger)\ x\le u_{2j+1}^+\lambda_0+u_{2j}^+\lambda_1 = u_{2j+1}^-\lambda_0+u_{2j}^+\lambda_1 \qquad\text{and}\qquad (\ddagger)\ -y\le u_{2j}^-\lambda_0+u_{2j-1}^-\lambda_1 = u_{2j}^-\lambda_0+u_{2j-1}^+\lambda_1. \]
    We inductively compute the region $\phi_{2j}^{-1}\cC_{2j}(\phi_{2j}\lambda)$, using Lemma~\ref{le:transformed inequalities} and the equality $\phi_{2j}^{-1}=\phi_2^{-j}$.

@@

    For $k=2j+1$, $j\ge0$, and $\lambda\in\cI$, we get $\cC_{2j+1}(\phi_{2j+1}\lambda)\subset\RR^2$ is given by the inequalities 
    \[ (\dagger')\ -x\le u_{2j+1}^+\lambda_0+u_{2j}^+\lambda_1 \qquad\text{and}\qquad (\ddagger')\ y\le u_{2j+2}^-\lambda_0+u_{2j+1}^-\lambda_1.\]
    Using that $\phi_{2j+1}^{-1}=\phi_2^{-j}\phi_1^{-1}$, we compute the image inductively as above.
    From \eqref{eq:forward mutation 1} and Lemma~\ref{le:imaginary stability}, we see that the boundary ray for $\cC_{2j+1}(\phi_{2j+1}\lambda)$ corresponding to ($\dagger'$) lies entirely in the region in which $\phi_1^{-1}$ acts according to the matrix $\left[ \begin{array}{cc} -1 & 0\\ c & 1 \end{array}\right]$ of determinant $-1$.
    Thus applying Lemma~\ref{le:transformed inequalities}, the inequality ($\dagger'$) is transformed by $\phi_1^{-1}$ into the inequality $x\le u_{2j+1}^+\lambda_0+u_{2j}^+\lambda_1$.
    The boundary ray corresponding to ($\ddagger'$) intersects both domains of linearity for $\phi_1^{-1}$ and thus produces the inequalities
    \[ cx+y\le u_{2j+2}^+\lambda_0+u_{2j+1}^+\lambda_1 \qquad\text{and}\qquad y\le u_{2j+2}^+\lambda_0+u_{2j+1}^+\lambda_1.\]

@@

  \begin{lemma}
    For $k \le -1$ and $\lambda\in\cI$, the region $\phi_k^{-1}\cC_k(\phi_k\lambda)\subset\RR^2$ is the polytope determined by the following inequalities:
    \begin{align}
      \label{ineq:1} u_{k+1}^-\mu_0+u_k^+\mu_1 &\le u_{k+1}^-\lambda_0+u_k^+\lambda_1;\\
      \label{ineq:2} u_k^-\mu_0+u_{k-1}^+\mu_1 &\le u_k^-\lambda_0+u_{k-1}^+\lambda_1;\\
      \label{ineq:3} u_k^-\mu_0-u_{k+1}^+\mu_1 &\le u_k^-\lambda_0+u_{k-1}^+\lambda_1;\\
      \label{ineq:4} -u_{k+2}^-\mu_0-u_{k+1}^+\mu_1 &\le u_k^-\lambda_0+u_{k-1}^+\lambda_1.
    \end{align}

    For $k=2j$, $j<0$, and $\lambda\in\cI$, we observe that $\cC_k(\phi_k\lambda)\subset\RR^2$ is given by the inequalities 
    \[ (\dagger)\ x\le u_{2j+1}^+\lambda_0+u_{2j}^+\lambda_1 \qquad\text{and}\qquad (\ddagger)\ -y\le u_{2j}^-\lambda_0+u_{2j-1}^-\lambda_1. \]
    We inductively compute the region $\phi_{2j}^{-1}\cC_{2j}(\phi_{2j}\lambda)$, using Lemma~\ref{le:transformed inequalities} and the equality $\phi_{2j}^{-1}=\phi_2^{-j}$.

    \subsubsection*{Claim:} For $i<0$, $\phi_{2i}^{-1}\cC_{2j}(\phi_{2j}\lambda)$ is the region determined by the inequalities 
    \begin{align*}
      \tag{a} u_{2i+1}^-x+u_{2i}^+y &\le u_{2j+1}^-\lambda_0+u_{2j}^+\lambda_1;\\
      \tag{b} u_{2i}^-x+u_{2i-1}^+y &\le u_{2j}^-\lambda_0+u_{2j-1}^+\lambda_1;\\
      \tag{c} u_{2i}^-x-u_{2i+1}^+y &\le u_{2j}^-\lambda_0+u_{2j-1}^+\lambda_1;\\
      \tag{d} -u_{2i+2}^-x-u_{2i+1}^+y &\le u_{2j}^-\lambda_0+u_{2j-1}^+\lambda_1.
    \end{align*}

    We see from \eqref{eq:two step mutation}, that $\lambda\in\cI$ implies the boundary ray for $\cC_{2j}(\phi_{2j}\lambda)$ corresponding to ($\dagger$) lies entirely in the region in which $\phi_2$ acts according to the matrix $\left[ \begin{array}{cc} bc-1 & b\\ -c & -1 \end{array}\right]$ of determinant 1.
    By Lemma~\ref{le:transformed inequalities}, the inequality ($\dagger$) transforms into the inequality $-x-by\le u_{2j+1}^+\lambda_0+u_{2j}^+\lambda_1$ which corresponds to (a) with $i=-1$.
    Similarly, the boundary ray for $\cC_{2j}(\phi_{2j}\lambda)$ corresponding to ($\ddagger$) intersects the three domains of linearity in which $\phi_2$ acts according to the matrices $\left[ \begin{array}{cc} bc-1 & b\\ -c & -1 \end{array}\right]$, $\left[ \begin{array}{cc} -1 & 0\\ -c & -1 \end{array}\right]$, $\left[ \begin{array}{cc} -1 & 0\\ 0 & -1 \end{array}\right]$, each of determinant 1.
    By Lemma~\ref{le:transformed inequalities}, the inequality ($\ddagger$) can be seen to transform by these into each of inequalities (b), (c), (d) with $i=-1$.
    \[ -cx-(bc-1)y\le u_{2j}^-\lambda_0+u_{2j-1}^-\lambda_1. \]
    \[ -cx+y\le u_{2j}^-\lambda_0+u_{2j-1}^-\lambda_1. \]
    \[ y\le u_{2j}^-\lambda_0+u_{2j-1}^-\lambda_1. \]
    This establishes the base of a downward induction on $i$.

    Assuming the inequalities (a)-(d) hold for $i$, we apply Lemma~\ref{le:transformed inequalities} with $\phi_2$.
    Both of the boundary rays corresponding to the inequalities (a) and (d) lies entirely in the region where $\phi_2$ acts according to the matrix $\left[ \begin{array}{cc} -1 & -b\\ c & bc-1 \end{array}\right]$, also the boundary segment corresponding to (b) intersects this region.
    Thus applying Lemma~\ref{le:transformed inequalities} to (a) gives the inequality 
    \[u_{2i+2}^-x+u_{2i+1}^+y=(u_3^+u_{2i}^--u_2^-u_{2i-1}^+)x+(u_2^+u_{2i}^--u_1^-u_{2i-1}^+)y\le u_{2j}^-\lambda_0+u_{2j-1}^+\lambda_1,\]
    which is the inequality (a) for $i+1$ by Lemma~\ref{eq:long Chebyshev recursion}; while applying this to (d) gives the inequality 
    \[-u_{2i+1}^-x-u_{2i}^+y=(-u_3^+u_{2i-1}^-+u_2^-u_{2i-2}^+)x+(-u_2^+u_{2i-1}^-+u_1^-u_{2i-2}^+)y\le u_{2j+1}^-\lambda_0+u_{2j}^+\lambda_1,\]
    which is the inequality (d) for $i+1$ again by Lemma~\ref{eq:long Chebyshev recursion}; finally applying this to (b) gives the inequality 
    \[u_{2i+3}^-x+u_{2i+2}^+y=(u_3^+u_{2i+1}^--u_2^-u_{2i}^+)x+(u_2^+u_{2i+1}^--u_1^-u_{2i}^+)y\le u_{2j+1}^-\lambda_0+u_{2j}^+\lambda_1,\]
    which is the inequality (b) for $i+1$.
    Similarly, the boundary segment corresponding to (c) lies entirely in the region where $\phi_2^{-1}$ acts according to the matrix $\left[ \begin{array}{cc} -1 & 0\\ c & -1 \end{array}\right]$.
    Thus applying Lemma~\ref{le:transformed inequalities} to (c) gives the inequality 
    \[-u_{2i+1}^-x-u_{2i}^+y=(u_1^+u_{2i-1}^--u_2^-u_{2i}^+)x+(-u_0^+u_{2i-1}^--u_1^-u_{2i}^+)y\le u_{2j+1}^-\lambda_0+u_{2j}^+\lambda_1,\]
    which is the inequality (d) for $i+1$ by Lemma~\ref{eq:long Chebyshev recursion}, in particular we see that the segment determined by (c) and the ray determined by (d) align in the image.
    Lastly, the boundary segment corresponding to (b) also intersects the regions where $\phi_2^{-1}$ acts according to the matrices $\left[ \begin{array}{cc} -1 & -b\\ 0 & -1 \end{array}\right]$ and $\left[ \begin{array}{cc} -1 & 0\\ 0 & -1 \end{array}\right]$ respectively.
    Applying Lemma~\ref{le:transformed inequalities} to (b) with the first matrix gives the inequality 
    \[-u_{2i+1}^-x+u_{2i+2}^+y=(-u_1^+u_{2i+1}^-+u_0^-u_{2i}^+)x+(u_2^+u_{2i+1}^--u_1^-u_{2i,+})y\le u_{2j+1}^-\lambda_0+u_{2j}^+\lambda_1,\]
    which is the inequality (c) for $i+1$ by Lemma~\ref{eq:long Chebyshev recursion}, while applying Lemma~\ref{le:transformed inequalities} to (b) with the second matrix gives the inequality 
    \[-u_{2i+1}^-x-u_{2i}^+y\le u_{2j+1}^-\lambda_0+u_{2j}^+\lambda_1,\]
    which again reproduces the inequality (d) and aligns with the previous segment and ray in the image.
    This completes the induction on $i$, proving the Claim and the result for $k$ even.

    For $k=2j-1$, $j \le 0$, and $\lambda\in\cI$, we get $\cC_{2j+1}(\phi_{2j+1}\lambda)\subset\RR^2$ is given by the inequalities 
    \[ (\dagger')\ -x\le u_{2j-1}^+\lambda_0+u_{2j-2}^+\lambda_1 = u_{2j-1}^-\lambda_0+u_{2j-2}^+\lambda_1 \qquad\text{and}\qquad (\ddagger')\ y\le u_{2j}^-\lambda_0+u_{2j-1}^-\lambda_1 = u_{2j}^-\lambda_0+u_{2j-1}^+\lambda_1.\]
    Using that $\phi_{2j-1}^{-1}=\phi_2^{-j}\phi_{-1}^{-1}$, we compute the image inductively as above.
    From \eqref{eq:forward mutation 1} and Lemma~\ref{le:imaginary stability}, we see that the boundary ray for $\cC_{2j+1}(\phi_{2j+1}\lambda)$ corresponding to ($\ddagger'$) lies entirely in the region in which $\phi_{-1}^{-1}$ acts according to the matrix $\left[ \begin{array}{cc} -1 & 0\\ 0 & -1 \end{array}\right]$ of determinant $-1$.
    Thus applying Lemma~\ref{le:transformed inequalities}, the inequality ($\ddagger'$) is transformed by $\phi_{-1}^{-1}$ into the inequality $-y\le u_{2j}^-\lambda_0+u_{2j-1}^+\lambda_1$.
    The boundary ray corresponding to ($\dagger'$) intersects both domains of linearity for $\phi_{-1}^{-1}$ and thus produces the inequalities
    \[ -x-by\le u_{2j-1}^-\lambda_0+u_{2j-2}^+\lambda_1 \qquad\text{and}\qquad -x\le u_{2j-1}^+\lambda_0+u_{2j-2}^+\lambda_1.\]

    \subsubsection*{Claim:} For $i\le 0$, $\phi_{2i-1}^{-1}\cC_{2j+1}(\phi_{2j+1}\lambda)$ is the region determined by the inequalities 
    \begin{align*}
      \tag{a$'$} u_{2i+1}^-x+u_{2i}^+y &\le u_{2j+1}^-\lambda_0+u_{2j}^+\lambda_1;\\
      \tag{b$'$} u_{2i+2}^-x+u_{2i+1}^+y &\le u_{2j+2}^-\lambda_0+u_{2j+1}^+\lambda_1;\\
      \tag{c$'$} -u_{2i}^-x+u_{2i+1}^+y &\le u_{2j+2}^-\lambda_0+u_{2j+1}^+\lambda_1;\\
      \tag{d$'$} -u_{2i}^-x-u_{2i-1}^+y &\le u_{2j+2}^-\lambda_0+u_{2j+1}^+\lambda_1.
    \end{align*}
    By essentially the same calculations as above, these inequalities reproduce under the action of $\phi_2^{-1}$ and this completes the proof.
  \end{proof}
  The following consequence is easily seen from the Chebyshev recursion.
  \begin{corollary}
    \begin{align*}
      u_k^-(x-\lambda_0)+u_{k-1}^+(y-\lambda_1) &\le 0;\\
      u_{k+1}^-(x-\lambda_0)+u_k^+(y-\lambda_1) &\le 0;\\
      -u_{k-1}^-(x-\lambda_0)+u_k^+(y-\lambda_1) &\le cu_k^+\lambda_0;\\
      -u_{k-1}^-(x-\lambda_0)-u_{k-2}^+(y-\lambda_1) &\le cu_k^+\lambda_0+bu_{k-1}^-\lambda_1.
    \end{align*}
  \end{corollary}
